Symmetric Circuits for Rank Gates

### 2017-11-16 sections/translating-formulas-to-cicuits.tex
13-Nov-2017
16 Nov

The algorithm \emph{merge} takes a circuit and an appropriate set of gates and
produces a new circuit by combining those gates into a single gate.

\begin{definition}
  The algorithm \emph{merge} takes as input a circuit $C = \langle G, \Omega,
  \Sigma , \Lambda, L \rangle$ and a set of syntactically equivalent gates $H =
  \{h_1, \ldots , h_p\}$ and outputs a new circuit $C' = \langle G', \Omega',
  \Sigma', \Lambda', L' \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_q\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $\vec{a} \in Ind(g)$ if $L(g) (\vec{a}) \in H$
    then $L'(g)(\vec{a}) = h_1$ and if $L(g)(\vec{a}) \notin H$ then
    $L'(g)(\vec{a}) = L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that `merge' can be implemented so as to run in time polynomial
in the size of the input circuit. We note that if the input circuit to `merge'
is symmetric (resp. has unique labels) then the output circuit will also be
symmetric (resp. have unique labels). The input and output circuits compute the
same function.

@@

The algorithm \emph{separate-labels} takes a circuit and outputs an equivalent
circuit such that every symmetric-gate has unique labels.

\begin{definition}
  The algorithm \emph{separate-labels} takes as input a $(\mathbb{B},
  \tau)$-circuit $C := \langle G, \Omega, \Sigma , \Lambda, L \rangle$ and
  outputs circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L' \rangle$. We
  now describe how these values are derived. We have that $\Omega' = \Omega'$
  and $\Lambda' = \Lambda$. We start with $G' = G$, $\Sigma' = \Sigma$ and $L' =
  L$ and updates their definitions iteratively according to the following
  procedure.

  Let $k$ denote a variable initialised to $1$. Let $S$ be the set of all gates
  of height $k$ that are symmetric-gates. For each $h_1 \in \bigcup_{g \in
    S}H_g$ let $g \in S$ be the first gate (using the order of the circuit
  derived from its encoding) such that $\vert L(g)^{-1}(h_1) \vert$ is maximal.
  Let $\{a_1, \ldots, a_p\} := L(g)^{-1}(h_1)$ (again, making use of the order).
  We add to $G'$ the gates $\{h_2, \ldots, h_p\}$ and for each $i \in \{2,
  \ldots, p\}$ we set $\Sigma'(h_i)$ such that $h_i$ is a $\land$ gate with a
  single output and $L'(h_i)$ such that $h_i$ takes only $h_{i-1}$ as an input.
  For each $g' \in S$ such that $h_1 \in H_g$ let $\{ b_1, \ldots, b_q \} :=
  L(g)^{-1}(h_1)$. For each $i \in [q]$ let $L'(g')(b_i) = h_i$. Iterate $k$ and
  repeat this procedure until $k$ exceeds the depth of the circuit.
\end{definition}

It can be shown that `separate-labels' can be implemented so as to run in time
polynomial in the size of the input circuit. We note that both the input and
output circuits compute the same function. Moreover, if all of the non-symmetric
gates in the input circuit have unique labels then the output circuit will have
unique labels. If the input circuit is symmetric and has the property that every
non-symmetric gate in the circuit has unique labels then the output circuit is
symmetric.

@@

The algorithm \emph{insert} takes as input two circuits and creates a new
circuit formed by inserting the gates of the first circuit into the second
circuit.

\begin{definition}
  The algorithm \emph{insert} takes as input a $(\mathbb{B}_1, \tau_1)$-circuit
  $C^1 := \langle G_1, \Omega_1, \Sigma_1 , \Lambda_1, L_1 \rangle$ and a
  $(\mathbb{B}_2, \tau_2)$-circuit $C^2 := \langle G_2, \Omega_2, \Sigma_2 ,
  \Lambda_2, L_2 \rangle$ and outputs a $(\mathbb{B}_1 \cup \mathbb{B}_2, \tau_1
  \uplus \tau_2)$-circuit $C = \langle G, \Omega, \Sigma , \Lambda, L \rangle$
  such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G = G_1 \uplus G_2$,
  \item $\Omega = \Omega_2$,
  \item $\Sigma = \Sigma_1 \uplus \Sigma_2$,
  \item $\Lambda = \Lambda_1 \uplus \Lambda_2$, and
  \item $L = L_1 \uplus L_2$.
  \end{itemize}
\end{definition}

It can be shown that `insert' can be implemented so as to run in time polynomial
in the size of the input circuit. If the input circuits are symmetric then the
output circuit will be symmetric.

The operation \emph{replace} removes a gate from a circuit and replaces it with
another gate.

\begin{definition}
  Formally, \emph{replace} is an algorithm that takes as input a circuit $C =
  \langle G, \Omega, \Sigma, \Lambda, L\rangle$ and a pair of gates $(g_1, g_2)$
  in $C$ and outputs a circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$ such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{g_1\}$,
  \item $\Omega'$ is an function from the domain of $\Omega$ to $G'$ such that
    for each $\vec{a} \in \dom(Omega)$ if $\Omega(\vec{a}) = g_1$ then
    $\Omega'(\vec{a}) = g_2$ and if then $\Omega'(\vec{a}) = \Omega(\vec{a})$.
  \item $\Sigma' = \restr{\Sigma}{G'}$
  \item $\Lambda' = \restr{\Lambda}{G'}$
  \item For each $g \in G'$ and each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) =
    g_1$ then $L'(g)(\vec{a}) = g_2$ and if not then $L'(g)(\vec{a}) =
    L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that `replace' can be implemented so as to run in time
polynomial in the size of the input circuit.

The algorithm \emph{tag} takes in a circuit, a gate in that circuit, a relation
symbol and a tuple and outputs a circuit computing but with an additional
gadget, constructed using the gates labelled by the given relation symbol and
tuple, marking the designated gate. The idea is that marking a gate can be used
as a step to ensure unique labels in a circuit.

\begin{definition}
  The algorithm \emph{tag} takes in a $(\mathbb{B}, \tau)$-circuit $C_n :=
  \langle G, \Omega, \Sigma, \Lambda, L \rangle$, a relation symbol $R \in
  \tau$, a gate $g \in G$ and a tuple $\vec{a} \in [n]^k$ for some $k$ and
  outputs the circuit $C_n' = \langle G', \Omega', \Sigma', \Lambda', L'\rangle$
  defined by the following algorithm. Let $G' = G \uplus \{\land_g, \land_{g'},
  \land_1, \ldots , \land_k\}$, $\Omega' = \Omega$, and $\Lambda' = \Lambda$.
  Define $L'$ and $\Sigma'$ as follows. For each $g \in G$ $\Sigma'(g) =
  \Sigma(g)$. For each $g \in \{\land_0, \land_1, \ldots , \land_k \}$,
  $\Sigma(g)$ is an $\AND$-gate that takes in two inputs. $\Sigma(\land_g)$ is
  an $\AND$-gate that takes in a single input. Set $g$ to be the lone input to
  $\land_g$. Set the two inputs to $\land_0$ to be $\land_g$ and $\land_1$. For
  each $i \in [k-1]$ set the inputs of $\land_i$ to be $\land_{i+1}$ and the
  gate $g_i \in G'$ such that $\Sigma(g_i) = R$ and $\Lambda^{-1}(g_i) =
  \vec{a}_i$, where $\vec{a}_i$ is the $k$-tuple $(\vec{a}(i), \ldots,
  \vec{a}(i))$. Set the inputs of $\land_k$ to be the constant gate $1$ and the
  gate $g_k$ such that $\Sigma(g) = R$ and $\Lambda^{-1}(g_k) = \vec{a}_k$.

@@

It can be shown that `tag' can be implemented so as to run in time polynomial in
the size of the input. The input and output circuits compute the same function.
Moreover, if the input circuit had unique labels then the output circuit will
have unique labels. In the event that the input relation symbol is obvious from
the context we omit it when invoking this algorithm.

The algorithm \emph{glue-circuits} takes in a designated circuit and another set
of appropriate circuits and forms a new circuit by replacing the relational
gates of the designated circuit with the output gates of the circuits in the
given set.

\begin{definition}
  The algorithm \emph{glue-circuits} takes as input a set of relation symbols
  $\tau_R := \{R_1, \ldots, R_t\}$ and $t+1$ circuits $\{C^0, \ldots, C^{t}\}$,
  where $C^0 = \langle G_0, \Omega_0, \Sigma_0, \Lambda_0, L_0 \rangle$ is a
  $(\mathbb{B}, \tau)$-circuit with $\tau_R \subseteq \tau$ and for each $i \in
  [t]$, $C^i = \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is a
  $(\mathbb{B}_i, \tau_i)$-circuit computing a query of arity $\arty(R_i)$. The
  algorithm outputs, for a given input, a circuit $C = \langle G, \Omega,
  \Sigma, \Lambda, L\rangle$ constructed as follows. Let $C^0' := C^0$. For each
  $i \in [t]$ fix a relation symbol $V_i \in \tau_i \setminnus \tau_R$. For each
  $i \in [t]$ construct the circuit $C^{i}' = \langle G_i', \Omega_i',
  \Sigma_i', \Lambda_i', L_i' \rangle $ by `inserting' $C^i$ into $C^{i-1}'$,
  and call the resultant circuit $D^{i-1}'$. For each relational gate $g$ in
  $D^{i-1}$ labelled by the symbol $R_i$, `replace' $g$ with the gate
  $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ and `tag' the gate
  $\Omega_i(\Lambda_{i-1}'(g))$ in $D^{i-1}$ using the tuple $\Lambda_{i-1}'(g)$
  and the relation symbol $V_i$. Output the circuit $C := C^t'$.
\end{definition}

It can be shown that `glue-circuits' can be implemented so as to run in time
polynomial in the combined size of the input circuits. Moreover, if the input
circuits are all symmetric (resp. have unique labels) then the output circuit
will be symmetric (resp. have unique labels). We sometimes refer to `gluing'
$(C^1, \ldots, C^t)$ into $C^0$ along $(R_1 , \ldots , R_t)$ when invoking
`glue-circuit' on the input $(C^0, \ldots , C^t)$ and $\tau_R = \{R_1, \ldots ,
R_t\}$.

@@

\begin{prop}
  There is an algorithm that takes in a $(\mathbb{B}, \tau)$-circuit with unique
  labels and outputs the syntactic equivalence relation on its gates. This
  algorithm runs in polynomial time.
  \label{prop:unique-labels-syntactic-equiv}
\end{prop}
\begin{proof}
\end{proof}

@@

  % it follows from Lemma \ref{}, that the syntactic equivalence relation on
  % $G_1 \uplus G_2$ can be efficiently computed.

>>>>

We first define the operation $\op{merge}$ that takes in a circuit and a set of
syntactically equivalent gates and produces a new circuit by combining those
gates into a single gate.

\begin{definition}
  The operation $\op{merge}(C, H)$ takes as input a circuit $C = \langle G,
  \Omega, \Sigma , \Lambda, L \rangle$ and a set of syntactically equivalent
  gates $H = \{h_1, \ldots , h_p\}$ and outputs the circuit $C' = \langle G',
  \Omega', \Sigma', \Lambda', L' \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_q\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $\vec{a} \in Ind(g)$ if $L(g) (\vec{a}) \in H$
    then $L'(g)(\vec{a}) = h_1$ and if $L(g)(\vec{a}) \notin H$ then
    $L'(g)(\vec{a}) = L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that $\op{merge}$ can be computed by an algorithm running in
time polynomial in the size of the input circuit. We note that the input and
output circuits compute the same function. Moreover, if the input circuit is
symmetric (resp. has unique labels) then the output circuit will also be
symmetric (resp. have unique labels).

The the operation $\op{merge-all}$ simply applies $\op{merge}$ to the syntactic
equivalence classes of some input set of gates.

\begin{definition}
  The operation $\op{merge-all}(C, H)$ takes as input an ordered circuit $C$ and
  a set of gates $H$ with unique labels and outputs another circuit $C'$. Let
  $H_1, \ldots, H_p$ be the equivalence classes of $H$ linearly ordered using
  the order on the circuit but such that for all $i, j \in [p]$ if $i \leq j$
  then the height of the gates in $H_i$ is less than or equal to the height of
  the gates in $H_j$. Let $C'$ be the result of successively executing
  $\op{merge}$ on each partition.
\end{definition}

For an input tuple $(C, H)$ it follows from Lemma \cite{} that the syntactic
equivalence relation on $H$ can be computed in time polynomial in the size of
$C$. Thus it can be shown that the operation $\op{merge-all}$ can be computed in
time polynomial in the size of the input circuit. As in the case of
$\op{merge}$, the operation $\op{merge-all}$ preserves symmetry, unique labels
and the ensures that the input and output circuits compute the same function. If
$C$ is an ordered circuit and $H$ is a set of gates with unique labels, then if
$g$ is a gate in both $C$ and $C' := \op{merge-all}(C, H)$, then if all of the input
gates of $g$ in $C$ are in $H$ then no two distinct input gates of $g$ in $C'$
are syntactically equivalent (but $g$ does not have unique labels in $C'$ as
$L'(g)$ may not be injective).

@@

The operation $\op{make-injective}$ takes in a circuit and a symmetric gate and
outputs an equivalent circuit but such that the selected gate now has injective
labels.

\begin{definition}
  The operation $\op{make-injective}(C, h)$ takes as input an ordered
  $(\mathbb{B}, \tau)$-circuit $C := \langle G, \Omega, \Sigma , \Lambda, L
  \rangle$ and a gate $h \in G$ and outputs the circuit $C' := \langle G',
  \Omega', \Sigma', \Lambda', L' \rangle$. Let $c = \max \{\vert
  L(g)^{-1}(h)\vert : g \in W(h, \cdot)\}$. If $c = 1$ then $C' := C$. If $c >
  1$ then $C'$ is defined such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \uplus \{h_2, \ldots, h_{c-1}\}$,
  \item $\Omega' = \Omega$,
  \item For each $g \in G$ let $\Sigma' (g) = \Sigma(g)$, otherwise let
    $\Sigma'(g)$ be a single-input $\AND$-gate.
  \item $\Lambda'= \Lambda$
  \item For each $g \in G'$ if $g \in G \setminus W(h, \cdot)$ then $L'(g) =
    L(g)$. If $g \in W(h, \cdot)$ let $\{\vec{a}_1, \ldots, \vec{a}_k\} :=
    L(g)^{-1}(h)$ and for each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) \neq h$
    then $L'(g)(\vec{a}) = L(g)(\vec{a})$ and if $\vec{a} = \vec{a_i} \in
    L(g)^{-1}(h)$ then let $L'(g)(\vec{a}) = h_i$.
  \end{itemize}
\end{definition}

It can be shown that $\op{make-injective}$ can be computed by an algorithm
running in time polynomial in the size of the input circuit. Moreover, the input
and output circuits of this operation compute the same function.

% Furthermore, $\op{make-injective}(C, h) = C'$ and there exists a symmetric
% gate $g \in C$ such that $\vert L(g)^{-1}(h) \vert > 1$ then the number of
% non-injective symmetric gates in $C'$ is strictly less than the number of
% non-injective symmetric gates in $C'$.

The operation $\op{make-injective-all}$ takes in a circuit and outputs an
equivalent circuit such that every symmetric-gate in that circuit has injective
labels.

\begin{definition}
  The operation $\op{make-injective-all}(C)$ takes as input an ordered
  $(\mathbb{B}, \tau)$-circuit $C := \langle G, \Omega, \Sigma , \Lambda, L
  \rangle$ and outputs the circuit $C' := \langle G', \Omega', \Sigma',
  \Lambda', L' \rangle$ defined as follows. Let $F(C*) = \op{make-injective}(C*,
  h)$, where $h$ is the first gate in $C*$ such that there exists a gate $g$ in
  $C*$ where $\vert L*(g)^{-1}(h)\vert > 1$. Let $(C^0, C^1, \ldots)$ be a
  sequence of circuits where $C^0 := C$ and for $i > 1$, $C^i = F(C^{i-1})$. It
  can be shown that this sequence converges after at most $\vert C \vert$ many
  steps. Let $C'$ be this limit.
\end{definition}

Since $\op{make-injective}$ can be computed in polynomial time, and the number
of iterations required in order to reach is a fixed-point in the definition of
$\op{make-injective-all}$ is bounded by the size of the input circuit, it
follows that $\op{make-injective-all}$ can be computed in time polynomial in the
size of the input circuit. Moreover, $\op{make-injective-all}$ preserves
symmetry, rigidity and unique labels. Moreover, the input and output circuits
compute the same function.

% $ $g$ in $C$ such that $W(h,g)$, let $g_{h_1}$ be the first gate in the
% circuit such that for all $g' \in S_j$, $\vert L(g_{h_1})^{-1}(h_1) \vert \geq
% L(g')^{-1}(h_1)$

% We define $C'$ as follows. Let $F$ be a function that takes in an ordered
% circuit $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ and outputs a
% circuit $C* := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ defined as
% follows. If for all $g \in G$, $L(g)$ is an injection then $F(C) = C$ and if
% not then $F(C)$ is defined as follows.

% Let $j \in [\depth(C)]$, and let ${G_0}' = G$, ${\Sigma_0}' = \Sigma$ and
% ${L}_0' = L$.

% Let $S_j$ be the set of all gates of height $j$ that are symmetric-gates. For
% each $h_1 \in \bigcup_{g \in S_j} H_g$ let $g_{h_1} \in S_j$ be the first gate
% in the circuit such that for all $g' \in S_j$, $\vert L(g_{h_1})^{-1}(h_1)
% \vert \geq L(g')^{-1}(h_1)$. Let $(a_1, \ldots, a_p) := L(g_{h_1})^{-1}(h_1)$
% (ordered according the to order on the circuit). We add to $G'$ the gates
% $\{h_2, \ldots, h_p\}$ and for each $i \in \{2, \ldots, p\}$ we set
% $\Sigma'(h_i)$ such that $h_i$ is a $\land$ gate with a single output and
% $L'(h_i)$ such that $h_i$ takes only $h_{i-1}$ as an input. For each $g' \in
% S$ such that $h_1 \in H_g$ let $\{ b_1, \ldots, b_q \} := L(g)^{-1}(h_1)$. For
% each $i \in [q]$ let $L'(g')(b_i) = h_i$. Iterate $k$ and repeat this
% procedure until $k$ exceeds the depth of the circuit.

% \begin{definition}
%   The operation \emph{separate-labels} takes as input an ordered $(\mathbb{B},
%   \tau)$-circuit $C := \langle G, \Omega, \Sigma , \Lambda, L \rangle$ and
%   outputs the circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L'
%   \rangle$. We define $C'$ as follows. Let $j \in [\depth(C)]$, and let
%   ${G_0}' = G$, ${\Sigma_0}' = \Sigma$ and ${L}_0' = L$. Let $S_j$ be the set
%   of all gates of height $j$ that are symmetric-gates. For each $h_1 \in
%   \bigcup_{g \in S_j} H_g$ let $g_{h_1} \in S_j$ be the first gate in the
%   circuit such that for all $g' \in S_j$, $\vert L(g_{h_1})^{-1}(h_1) \vert
%   \geq L(g')^{-1}(h_1)$. Let $(a_1, \ldots, a_p) := L(g_{h_1})^{-1}(h_1)$
%   (ordered according the to order on the circuit). We add to $G'$ the gates
%   $\{h_2, \ldots, h_p\}$ and for each $i \in \{2, \ldots, p\}$ we set
%   $\Sigma'(h_i)$ such that $h_i$ is a $\land$ gate with a single output and
%   $L'(h_i)$ such that $h_i$ takes only $h_{i-1}$ as an input. For each $g' \in
%   S$ such that $h_1 \in H_g$ let $\{ b_1, \ldots, b_q \} := L(g)^{-1}(h_1)$.
%   For each $i \in [q]$ let $L'(g')(b_i) = h_i$. Iterate $k$ and repeat this
%   procedure until $k$ exceeds the depth of the circuit.
% \end{definition}

@@

The operation $\op{insert}$ takes as input two circuits and outputs a circuit
formed by inserting the gates of the first circuit into the second circuit.

\begin{definition}
  The operation $\op{insert}(C^1, C^2)$ takes as input a $(\mathbb{B}_1,
  \tau_1)$-circuit $C^1 := \langle G_1, \Omega_1, \Sigma_1 , \Lambda_1, L_1
  \rangle$ and a $(\mathbb{B}_2, \tau_2)$-circuit $C^2 := \langle G_2, \Omega_2,
  \Sigma_2 , \Lambda_2, L_2 \rangle$ and outputs a $(\mathbb{B}_1 \cup
  \mathbb{B}_2, \tau_1 \uplus \tau_2)$-circuit $C = \langle G, \Omega, \Sigma ,
  \Lambda, L \rangle$ where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G = G_1 \uplus G_2$,
  \item $\Omega = \Omega_2$,
  \item $\Sigma = \Sigma_1 \uplus \Sigma_2$,
  \item $\Lambda = \Lambda_1 \uplus \Lambda_2$, and
  \item $L = L_1 \uplus L_2$.
  \end{itemize}
\end{definition}

It can be shown that $\op{insert}$ can be be computed in polynomial time. If the
input circuits are symmetric then the output circuit will be symmetric.

The operation $\op{replace}$ removes a gate from a circuit and replaces it with
another gate.

\begin{definition}
  The operation $\op{replace}(C, g_1, g_2)$ takes as input a circuit $C =
  \langle G, \Omega, \Sigma, \Lambda, L\rangle$ and a pair of gates $(g_1, g_2)$
  in $C$ and outputs a circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$ such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{g_1\}$,
  \item $\Omega'$ is an function from the domain of $\Omega$ to $G'$ such that
    for each $\vec{a} \in \dom(Omega)$ if $\Omega(\vec{a}) = g_1$ then
    $\Omega'(\vec{a}) = g_2$ and if then $\Omega'(\vec{a}) = \Omega(\vec{a})$.
  \item $\Sigma' = \restr{\Sigma}{G'}$
  \item $\Lambda' = \restr{\Lambda}{G'}$
  \item For each $g \in G'$ and each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) =
    g_1$ then $L'(g)(\vec{a}) = g_2$ and if not then $L'(g)(\vec{a}) =
    L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that the operation $\op{replace}$ can be computed in time
polynomial in the size of the input circuit.

The operation $\op{tag}$ takes in a circuit, a gate in that circuit, a relation
symbol and a tuple and outputs a circuit computing the same function but with an
additional gadget, constructed using the gates labelled by the given relation
symbol and tuple, marking the designated gate. The idea is that marking a gate
can be used as a step to ensure unique labels in a circuit.

\begin{definition}
  The operation $\op{tag}(C, R, g, \vec{a})$ takes in a $(\mathbb{B},
  \tau)$-circuit $C_n := \langle G, \Omega, \Sigma, \Lambda, L \rangle$, a
  relation symbol $R \in \tau$, a gate $g \in G$ and a tuple $\vec{a} \in [n]^k$
  for some $k$ and outputs the circuit $C_n' = \langle G', \Omega', \Sigma',
  \Lambda', L'\rangle$ defined by the following algorithm. Let $G' = G \uplus
  \{\land_g, \land_{g'}, \land_1, \ldots , \land_k\}$, $\Omega' = \Omega$, and
  $\Lambda' = \Lambda$. Define $L'$ and $\Sigma'$ as follows. For each $g \in G$
  $\Sigma'(g) = \Sigma(g)$. For each $g \in \{\land_0, \land_1, \ldots , \land_k
  \}$, $\Sigma(g)$ is an $\AND$-gate that takes in two inputs. $\Sigma(\land_g)$
  is an $\AND$-gate that takes in a single input. Set $g$ to be the lone input
  to $\land_g$. Set the two inputs to $\land_0$ to be $\land_g$ and $\land_1$.
  For each $i \in [k-1]$ set the inputs of $\land_i$ to be $\land_{i+1}$ and the
  gate $g_i \in G'$ such that $\Sigma(g_i) = R$ and $\Lambda^{-1}(g_i) =
  \vec{a}_i$, where $\vec{a}_i$ is the $k$-tuple $(\vec{a}(i), \ldots,
  \vec{a}(i))$. Set the inputs of $\land_k$ to be the constant gate $1$ and the
  gate $g_k$ such that $\Sigma(g) = R$ and $\Lambda^{-1}(g_k) = \vec{a}_k$.

@@

% \begin{definition}
%   The algorithm \emph{tag} takes in a $(\mathbb{B}, \tau)$-circuit $C_n :=
%   \langle G, \Omega, \Sigma, \Lambda, L \rangle$, a relation symbol $R \in
%   \tau$, a gate $g \in G$ and a tuple $\vec{a} \in [n]^k$ for some $k$ and
%   outputs the circuit $C_n' = \langle G', \Omega', \Sigma', \Lambda',
%   L'\rangle$ defined by the following algorithm. Let $G' = G \uplus \{\land_g,
%   \land_{g'}, \land_1, \ldots , \land_k\}$, $\Omega' = \Omega$, and $\Lambda'
%   = \Lambda$. Define $L'$ and $\Sigma'$ as follows. For each $g \in G$
%   $\Sigma'(g) = \Sigma(g)$. For each $g \in \{\land_0, \land_1, \ldots ,
%   \land_k \}$, $\Sigma(g)$ is an $\AND$-gate that takes in two inputs.
%   $\Sigma(\land_g)$ is an $\AND$-gate that takes in a single input. Set $g$ to
%   be the lone input to $\land_g$. Set the two inputs to $\land_0$ to be
%   $\land_g$ and $\land_1$. For each $i \in [k-1]$ set the inputs of $\land_i$
%   to be $\land_{i+1}$ and the gate $g_i \in G'$ such that $\Sigma(g_i) = R$
%   and $\Lambda^{-1}(g_i) = \vec{a}_i$, where $\vec{a}_i$ is the $k$-tuple
%   $(\vec{a}(i), \ldots, \vec{a}(i))$. Set the inputs of $\land_k$ to be the
%   constant gate $1$ and the gate $g_k$ such that $\Sigma(g) = R$ and
%   $\Lambda^{-1}(g_k) = \vec{a}_k$.

%   For each gate $h \in G \setminus \{g\}$ and $\vec{b} \in \ind(h)$, if
%   $L(h)(\vec{b}) = g$ then let $L'(h)(\vec{b}) = \land_0$, otherwise let
%   $L'(h)(\vec{b}) = L(h)(\vec{b})$.

% \end{definition}

It can be shown that $\op{concatenate}$ can be implemented so as to run in time
polynomial in the size of the input. The input and output circuits compute the
same function. Moreover, if the input circuit had unique labels then the output
circuit will have unique labels. In the event that the input relation symbol is
obvious from the context we omit it when invoking this algorithm.

The operation $\op{concatenate}$ takes in a circuit and a set of appropriate
circuits and forms a new circuit by replacing the relational gates of the
designated circuit with the output gates of the circuits in the given set of
circuits.

\begin{definition}
  The operation $\op{concatenate}(C, \mathcal{C}, \tau_R, R)$, where $C :=
  \langle G, \Omega, \Sigma, \Lambda, L \rangle$ is a $(\mathbb{B},
  \tau)$-circuit, $\tau_R := (R_1, \ldots, R_t)$ is a sequence of relation
  symbols such that $\tau_R \subseteq \tau$, $\mathcal{C} := ( C^1, \ldots,
  C^t)$ is a finite sequence of circuits, where for each $i \in [t]$, $C^i =
  \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is a $(\mathbb{B}_i,
  \tau_i)$-circuit such that $\tau_i \not\subset \tau_R$ and $C^i$ computes a
  query of arity $\arty(R_i)$, and $R$ is a relation symbol in $(\tau \setminus
  \tau_R) \cup (\bigcup_{1 \leq i \leq t}\tau_i))$. The operation outputs a
  circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L' \rangle$ defined from
  the input as follows. Note that all circuits take in structures of size $n$.

  Let ${C^0}' := \langle G^0, \Omega^0, \Sigma^0, \Lambda^0, L^0 \rangle := C$.
  For each $i \in [t]$ fix a relation symbol $V_i \in \tau_i \setminus \tau_R$.
  For each $i \in [t]$ construct the circuit ${C^{i}}' = \langle G_i',
  \Omega_i', \Sigma_i', \Lambda_i', L_i' \rangle $ by $\op{inserting}$ $C^i$
  into ${C^{i-1}}'$, and call the resultant circuit ${D^{i-1}}'$. For each
  relational gate $g$ in $D^{i-1}$ labelled by the symbol $R_i$, $\op{replace}$
  $g$ with the gate $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ and $\op{tag}$
  the gate $\Omega_i(\Lambda_{i-1}'(g))$ in $D^{i-1}$ using the tuple
  $\Lambda_{i-1}'(g)$ and the relation symbol $R$. Output the circuit $C' :=
  {C^t}'$.
\end{definition}

% \begin{definition}
%   The operation $\op{concatenate-circuits}(C, \mathcal{C}, \tau_R) = C'$,
%   where $C := \langle G, \Omega, \Sigma, \Lambda_0, L_0 \rangle$ is a
%   $(\mathbb{B}, \tau)$-circuit, $\mathcal{C}$ is a finite set of circuits
%   takes as input a set of relation symbols $\tau_R := \{R_1, \ldots, R_t\}$
%   and $t+1$ circuits $\{C^0, \ldots, C^{t}\}$, where $C^0 = \langle G_0,
%   \Omega_0, \Sigma_0, \Lambda_0, L_0 \rangle$ is a $(\mathbb{B},
%   \tau)$-circuit with $\tau_R \subseteq \tau$ and for each $i \in [t]$, $C^i =
%   \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is a
%   $(\mathbb{B}_i, \tau_i)$-circuit computing a query of arity $\arty(R_i)$.
%   The algorithm outputs, for a given input, a circuit $C = \langle G, \Omega,
%   \Sigma, \Lambda, L\rangle$ constructed as follows. Let ${C^0}' := C^0$. For
%   each $i \in [t]$ fix a relation symbol $V_i \in \tau_i \setminus \tau_R$.
%   For each $i \in [t]$ construct the circuit ${C^{i}}' = \langle G_i',
%   \Omega_i', \Sigma_i', \Lambda_i', L_i' \rangle $ by `inserting' $C^i$ into
%   ${C^{i-1}}'$, and call the resultant circuit ${D^{i-1}}'$. For each
%   relational gate $g$ in $D^{i-1}$ labelled by the symbol $R_i$, `replace' $g$
%   with the gate $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ and `tag' the
%   gate $\Omega_i(\Lambda_{i-1}'(g))$ in $D^{i-1}$ using the tuple
%   $\Lambda_{i-1}'(g)$ and the relation symbol $V_i$. Output the circuit $C :=
%   {C^t}'$.
% \end{definition}

It can be shown that $\op(concatenate)$ can be computed in time polynomial in
the combined size of the input circuits. Moreover, if the input circuits are all
symmetric (resp. have unique labels) then the output circuit will be symmetric
(resp. have unique labels). We sometimes refer to $\op(concatenate)$ $(C^1,
\ldots, C^t)$ into $C^0$ along $(R_1 , \ldots , R_t)$ when invoking
`glue-circuit' on the input $(C^0, \ldots , C^t)$ and $\tau_R = \{R_1, \ldots ,
R_t\}$.

@@

\begin{prop}
  Let $C_n := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
  \tau)$-circuit with unique labels. There is a deterministic algorithm that
  takes in such a circuit and outputs the syntactic equivalence relation on $G$.
  Moreover, this algorithm runs in time polynomial in the size of the circuit.
  \label{prop:unique-labels-syntactic-equiv}
\end{prop}
\begin{proof}
  We proceed to define a relation $\sim$ by induction on height (i.e. distance
  from an input gate). This relation will be the syntactic equivalence relation
  on $G$. It is easy to check the conditions for syntactic equivalence for gates
  of height 0 as all such gates are input gates and so are either constant gates
  or relational gates. Suppose $g$ and $h$ are gates of height greater than
  zero. We only need to check conditions (i), (iv) and (v) of syntactic
  equivalence. It is easy to check (i) and (v). If either condition is false
  then $g$ and $h$ are not syntacticly equivalent. So suppose $g$ and $h$ are
  two internal gates such that $\Sigma (g) = \Sigma(h)$ and if they are both
  output gates then $\Omega^{-1}(g) = \Omega^{-1}(h)$. It remains to check if
  $L(g)/{\sim}$ and $L(h)/{\sim}$ are isomorphism equivalent. First check if
  $H_h /{\sim} = H_g /{\sim}$. If not, $g$ and $h$ are not syntactically
  equivalent. So suppose $H_h /{\sim} = H_g /{\sim}$. We know that $C_n$ has
  unique labels, and so $L(g)/{\sim}$ and $L(h)/{\sim}$ are injections so are
  isomorphism equivalent if, and only if, $L(g)^{-1}L(h)$ acts like an
  isomorphism. This can be checked by first checking that the function induces a
  sorted permutation on the universe of $g$ (which can be done by iterating over
  the index of $g$). This is sufficient as $L(g)(L(g)^{-1}L(h)) = L(h)$.

  It is easy to see that, given a reasonable binary encoding of a circuit, the
  above procedure can be implemented so as to run in polynomial time, and that
  the relation $\sim$ is the syntactic equivalence relation on $C_n$.
\end{proof}

% \begin{prop}
%   There is an algorithm that takes in a $(\mathbb{B}, \tau)$-circuit $C$ with
%   unique labels and outputs the syntactic equivalence relation on its gates.
%   This algorithm runs in time polynomial in the size of $C$.
%   \label{prop:unique-labels-syntactic-equiv}
% \end{prop}
% \begin{proof}
%   We build of the relation by induction. Let $g$
% \end{proof}

@@

% it follows from Lemma \ref{}, that the syntactic equivalence relation on $G_1
% \uplus G_2$ can be efficiently computed.
